\def \root {..}			% path to root (/notes)
\input{\root/template/preamble.tex}

\title{Overview}
\author{Patrick Oare}
\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle

These notes are my attempt to put what I've learned of math and physics during my undergraduate and graduate degrees down onto paper. They'll be separated into Physics and Math directories. There will typically be some overlap in content between the note sets in each directory, but the physics notes will usually be written in a bit more of a free style, whereas the math notes will tend to be more precise. All notes should provide a good amount of examples that help reinforce the material, which I find to be especially helpful as the material gets more abstract. 

The notes in this repository are in a variety of different stages. Notes marked ``completed" have been edited and are more or less in their final form (although if you find an error, please let me know!). Notes marked ``in-progress" need to be edited and will likely be missing some content. Notes marked ``future" are planned for the future when I either have the time to write them, or actually get around to learning the material.

\section{Physics notes}

\subsection{Completed notes}

\begin{itemize}
	\item 
\end{itemize}

\subsection{In-progress notes}

\begin{itemize}
	\item The Standard Model.
	\item Instantons and solitons.
	\item Colliders.
	\item Hadron structure.
	\item Representation theory.
	\item The Lorentz and Poincar\'e groups.
\end{itemize}

\subsection{Future notes}

\begin{itemize}
	\item Renormalization.
	\item Lattice gauge theory.
	\item Topology in lattice gauge theory.
\end{itemize}

\section{Math notes}

\subsection{Completed notes}

\begin{itemize}
	\item Metric spaces.
	\item Measure theory. (finish a few proofs)
\end{itemize}

\subsection{In-progress notes}

\begin{itemize}
	\item Probability and stochastic processes.
	\item Point-set topology. (finish a few proofs)
	\item Differential forms.
	\item Complex analysis.
	\item Set theory.
	\item The fundamental group.
	\item Topological manifolds.
	\item Homology.
	\item Algebra (groups, rings, fields, modules, polynomials).
	\item Galois theory.
\end{itemize}

\subsection{Future notes}

\begin{itemize}
	\item Mathematical quantum mechanics.
	\item Differential manifolds.
	\item Differential geometry.
	\item Fiber bundles.
	\item Characteristic classes.
	\item Stochastic calculus.
	\item Random field theory.
	\item Mathematical statistics. 
	\item Functional analysis.
\end{itemize}

\end{document}